---
title: "GSEA_Joint_TCT_TNF"
author: "Bing Shui"
date: "4/19/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Prepare libraries
```{r prepare libraries}
suppressMessages(
  c(library(tidyverse))
)

```


## Load GSEA results
For this analysis, I would like to take a look at the GSEA pathways that are differentially enriched between MK2i and vehicle treated inflamed samples in TCT and TNF model to see how much overlap/difference we observe.

```{r loading dataset}
gsea_imki_iv_tct_down <- read_csv("~/OneDrive - Harvard University/Haigis Lab/Projects/Colloboration Projects/IBD RNA-Seq analysis/Analysis/TCT/GSEA/I_MKI vs I_V/Hallmark/gsea_report_for_0_1586820351136.csv")
gsea_imki_iv_tct_up <- read_csv("~/OneDrive - Harvard University/Haigis Lab/Projects/Colloboration Projects/IBD RNA-Seq analysis/Analysis/TCT/GSEA/I_MKI vs I_V/Hallmark/gsea_report_for_1_1586820351136.csv")
gsea_imki_iv_tct <- rbind(gsea_imki_iv_tct_down, gsea_imki_iv_tct_up)

gsea_imki_iv_tnf_down <- read_csv("~/OneDrive - Harvard University/Haigis Lab/Projects/Colloboration Projects/IBD RNA-Seq analysis/Analysis/TNF/GSEA/I_MKI vs I_V/Hallmark/gsea_report_for_0_1586821602749.csv")
gsea_imki_iv_tnf_up <- read_csv("~/OneDrive - Harvard University/Haigis Lab/Projects/Colloboration Projects/IBD RNA-Seq analysis/Analysis/TNF/GSEA/I_MKI vs I_V/Hallmark/gsea_report_for_1_1586821602749.csv")
gsea_imki_iv_tnf <- rbind(gsea_imki_iv_tnf_down, gsea_imki_iv_tnf_up)

```

## Cross analysis
First we filter the list and select only pathways that are significantly enriched using standard pval < 0.05 and q (padj) < 0.1.
```{r select significant pathways}
gsea_imki_iv_tct <- gsea_imki_iv_tct[gsea_imki_iv_tct$`NOM p-val` < 0.5 & gsea_imki_iv_tct$`FDR q-val` < 0.1, ]
gsea_imki_iv_tnf <- gsea_imki_iv_tnf[gsea_imki_iv_tnf$`NOM p-val` < 0.5 & gsea_imki_iv_tnf$`FDR q-val` < 0.1, ]
```

First we look at the overlap.
```{r overlapping pathways}
overlap <- intersect(gsea_imki_iv_tct$NAME, gsea_imki_iv_tnf$NAME)

# now let's check if all the overlapping pathways are enriched in the same direction
opposite <- c()
for (i in overlap) {
  tct_nes <- gsea_imki_iv_tct$NES[gsea_imki_iv_tct$NAME == i]
  tnf_nes <- gsea_imki_iv_tnf$NES[gsea_imki_iv_tnf$NAME == i]
  dir <- tct_nes * tnf_nes
  if (dir < 0) {
    print(paste("Pathway enriched in opposite directoin:" ,i))
    opposite <- c(opposite, i)
  }
}

print(paste("The number of overlapping enriched pathways is:", length(overlap)- length(opposite)))
overlap[!overlap%in%opposite]
```

Then we can take a look at uniquely enriched pathways.
 
```{r show unique pathways}
tct_unique <- setdiff(gsea_imki_iv_tct$NAME, overlap)
tnf_unique <- setdiff(gsea_imki_iv_tnf$NAME, overlap)
print(paste("These are the pathways uniquely enriched in TCT:", tct_unique))
print(paste("These are the pathways uniquely enriched in TNF:", tnf_unique))
```



